---
title: "My first R markdown"
subtitle: "Created fully by myself"
author: "Viktoria Konya"
date: "September 30, 2021"
output:
    prettydoc::html_pretty:
    theme: tactile
    highlight: github
    toc: true
    #toc_depth: 1

---
This is my first markdown.

# Aim

In this project I am practicing R markdown features.

I would like to:

1. understand R markdown
2. learn new things
3. create a nice document

# Theme
To change the theme of the document I am using Prettydocs.

# Codes
I include a code chunk with a comment in it.


```{r}
# I am the comment :D
```


I also say hello to the world with a print() and for that I am only showing the output of the code.

```{r}
## [1] "Hello World!"
```


# Tables
With kableExtra I show the first 5 rows of the ChickWeight dataset. For this I will use the kableExtra package which I imported in a separate chunk.

For the table I used the classic style.

```{r echo = F , warning = F, message = F}
# Import libraries
library(kableExtra)
```

```{r echo = F}
dt <- ChickWeight[1:5, ]
kable(dt) %>%
  kable_styling(full_width = F)
```

# Plots
I included one plot. It need to:

* be aligned to the center
* have a size of 70%
* hide the codes creating the plot
* hide warning message

```{r echo = F , warning = F, message = F}
# Import libraries
library(ggplot2)
```

```{r echo = F , warning = F, message = F}
ggplot(cars, aes(x=speed, y=dist)) +
  geom_point(color = "green4") +
  labs(title = "Chart of cars",
       x = "Speed",
       y = "Distance") +
  theme_bw()
```





