---
title: "My fantastic R markdown"
subtitle: "This is the cool subtitle of my document"
author: "Viki Mészáros"
date: "21/09/2021"
output: 
    html_document:
        toc: true
        code_download: true
        theme: flatly
---


# Write with Markdown

You can easily include text to a markdown file.
With two spaces  
you can start a new paragraph.

When you include a backslash\
you will have a new line.

You can also play around with the text you include. You can create **bold** or *italic* parts or could even ~~strikethrough~~ some words. You may need to include some superscrips x^2^ or subsripts x~1~.

---

# Headers

When creating documents you will most probably need to use some headings.

## Really big ones
### Smaller ones
##### Or even pretty smalls

---

# Lists

You can include cool ordered and unordered lists as well.

**Thing I like the most**

- Coding
- R
- The R Mentoring Class
  - Misi
  - Viki


*My aims for the year*

1. Become a pro in coding
2. Attend all R Mentoring classes
3. Enjoy my studies

-------------------

# Links

When you would like to put a link in your document, you can put the whole thing in the document, or you can also link it to a word which looks much better.

Here is a cool website to style my markdowns. <https://prettydoc.statr.me/>

Here is a cool [website](https://prettydoc.statr.me/) to style my markdowns.


---

# Equations

During your studies you will probably need to include some equations as well.

You can do this within your text. $e^{i \pi} + 1 = 0$ And continue explaining.

Or you may create a separate block.

$$E = mc^{2}$$

---

# Chunk options

After text elements the other most important thing you will put into your markdown is code.

You can include this between 

```{r chunk-name, include=F} 
#here comes your code
```


There are several different options you can use within the {}.
Without anything the code you write will be shown and the results as well.

```{r load data 1}

df = iris
head(df, 3)

```


With `eval=FALSE` (or `eval=F`) the code won't be evaluated, but it will be printed out.

```{r load data 2, eval=F}

df = iris
head(df, 3)

```


With `echo=F` you can only show the output of the chunk but hide the code you used to get it. This is mostly used for figures.

```{r load data 4, echo=F}

df = iris
head(df, 3)

```


With `include=F` you can exclude the chunk. It will run in the background. This is useful for data import or package loadings.

```{r load data 3, include=F}

df = iris
head(df, 3)

```

---

# Figures

For example here I am only showing the fig without the codes.

```{r iris fig, echo=F}
library(ggplot2)

p <- ggplot(data=iris,mapping=aes(x=Petal.Length,y=Petal.Width))+
  geom_point(aes(color=Sepal.Width))+
  geom_smooth(method="lm")+
  scale_color_viridis_c()+
  scale_x_continuous(breaks=1:8)+
  labs(title="This Is A Title",subtitle="This is a subtitle",x=" Petal Length", 
       y="Petal Width", caption="This is a little caption.")+
  facet_wrap(~Species)+
  theme_bw()+
  theme(
    axis.title=element_text(color="cyan4",face="bold", size = 11),
    plot.title=element_text(color="cyan4",face="bold", size = 18),
    plot.subtitle=element_text(color="cyan4", face = "italic", size = 14),
    panel.grid=element_blank(),
    legend.position = "none"
  )

show(p)
```

Here is the code for generate the plot
```{r, message=F}
library(ggplot2)

p <- ggplot(data=iris,mapping=aes(x=Petal.Length,y=Petal.Width))+
  geom_point(aes(color=Sepal.Width))+
  geom_smooth(method="lm")+
  scale_color_viridis_c()+
  scale_x_continuous(breaks=1:8)+
  labs(title="This Is A Title",subtitle="This is a subtitle",x=" Petal Length", 
       y="Petal Width", caption="This is a little caption.")+
  facet_wrap(~Species)+
  theme_bw()+
  theme(
    axis.title=element_text(color="cyan4",face="bold", size = 11),
    plot.title=element_text(color="cyan4",face="bold", size = 18),
    plot.subtitle=element_text(color="cyan4", face = "italic", size = 14),
    panel.grid=element_blank(),
    legend.position = "none"
  )

show(p)
```


There are two extra lines before the graph. It is also possible to hide them. For this you can use `warning=F` to hide warning messages (*for example*: Warning: package 'ggplot2' was built under R version 4.0.5)


```{r, echo=F, warning=F}
show(p)
```

With `message=F` you can also hide any additional messages. (*for example*: 'geom_smooth()' using formula 'y ~ x')

```{r, echo=F, message=F}
show(p)
```

---

You can also play around with the alignment of the figure.

```{r, echo=F, message=F, fig.align='center'}
show(p)
```

Or can change the size by the lines `out.width=` command. After the equal sign you can use percentages like *"50%"* or pixel sizes like *"300px"*

**Using the measure "50%"**

```{r, echo=F, message=F, fig.align='right', out.width="50%"}
show(p)
```

**Using the measure "500px"**

```{r, echo=F, message=F, fig.align='center', out.width="500px"}
show(p)
```




# Tables

There are some extra option to show your data in a markdown. The package I used the most was [KableExtra](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html).

This is what the basic table would look like.

```{r}
head(df, 3)
```

This is how it looks with kableExtra.

```{r}
library(kableExtra)

head(df, 3) %>% 
  kbl() 
```

You can add different styles.

```{r}
head(df, 3) %>% 
  kbl() %>% 
  kable_classic(full_width = F)
```

Or even some interactive features like hovering.

```{r, }
head(df, 3) %>% 
  kbl() %>% 
  kable_paper(c("striped", "hover"), full_width = F)
```


There are some additional packages you can use to format your tables nicely. 

- **flextable**
- **gt**
- **kable**


# Styles

To add some extra flavor to your docs you can use different themes. For this the easiest is to use [Bootswatch](https://bootswatch.com/). To change the styles you just need to replace *"flatly"* in the header to one of the names you can see on the website.

Another way is to use [Prettydoc](https://prettydoc.statr.me/). For this you can check their website and see how this works.








