---
title: "week-5-tasks"
author: "Viki-Misi"
date: "October 07, 2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Dplyr tasks

Load the `nycflights13` package and **write code**, to reproduce the same result that showed here while you anser the qustions!
```{r, message=FALSE}

# comment multiple lines 
# ctrl + shift + c

# install.packages('nycflights13')
library(nycflights13)
library( tidyverse )

```


Help
```{r}
# df %>%
#   filter() %>% 
#   group_by()%>%
#   summarise( ) %>%
#   arrange()%>%
#   head()
```




# Which airplane was in the air the most time?
```{r, echo=FALSE,message=F, warning=FALSE}
flights %>%
  filter(!is.na(tailnum)) %>% 
  group_by(tailnum)%>%
  summarise('in_air'= sum(air_time) ) %>%
  arrange(-in_air)%>%
  head(1)

```

# The best carriers based on arriving (on average the `arr_delay` is the least, it can be negativ), also show the number of flights
```{r, echo=FALSE, message=F, warning=FALSE}
# number of flights from origin
flights %>%
  group_by(carrier)%>%
  summarise('mean_arr_delay'=mean(arr_delay, na.rm = T), 
            'n_flights'= n()) %>%
  arrange(mean_arr_delay) %>% 
  head(1)



```

# What is the name of the arline company which had the most arrival delay on average in August

```{r, echo=FALSE, message=F, warning=FALSE}
t <- 
flights %>%
  filter(month==8) %>% 
  group_by(carrier)%>%
  summarise('mean_delay'=mean(arr_delay, na.rm = T)) %>% 
  arrange(-mean_delay) %>% 
  head(1) 
merge(t, airlines, by = 'carrier') %>% select(name)

```


Who is the manufacturer of the flights who flow the most times from `EWR` and how much times it flow?
```{r, echo=FALSE}
t <- 
flights %>% 
  filter(!is.na(tailnum),
         origin%in%c('EWR')) %>% 
  group_by(tailnum) %>% 
  summarise('flight_times'= n()) %>% 
  arrange(-flight_times) %>% 
  head(1)
merge(t,planes, by = 'tailnum' ) %>% select(tailnum, manufacturer, flight_times)

```


How many seats does the flights which flow tho most times from `JFK` to `MCO`?
```{r, echo=FALSE}
t <-
flights %>% 
  filter(origin=='JFK', dest=='MCO') %>% 
  group_by(tailnum) %>% 
  summarise('flight_times'=n()) %>%
  arrange(-flight_times) %>% 
  head(1)
merge(t,planes ) %>% select(tailnum,seats , flight_times)



```



# Ggplot tasks

Use the first 1000 rows of flights datasets 

```{r}
df <- flights %>% head(1000)
```

Reproduce the ggplots! 

```{r, echo=FALSE, warning=F, message=F, warning=FALSE}


# http://statseducation.com/Introduction-to-R/modules/graphics/cont/

#1
ggplot(df, aes(x=distance, y= arr_delay, color=carrier)) +
  geom_point()+
  labs(x='Arrival delay', y='Distance')
  

#2
ggplot(df, aes(x=distance, y= arr_delay, size=air_time)) +
  geom_point()+
  labs(x='Arrival delay', y='Distance')

#3
ggplot(df, aes(x=distance, y= arr_delay, shape = carrier)) +
  geom_point()+
  labs(x='Arrival delay', y='Distance')

 
ggplot(df, aes(x=distance, y= arr_delay)) +
  geom_point() +
  facet_wrap(~carrier)+
  labs(x='Arrival delay', y='Distance')
 
 
ggplot(df, aes(x=distance, y= arr_delay)) + 
  geom_point() +
  geom_smooth()+
  labs(x='Arrival delay', y='Distance')
 
ggplot(df, aes(x=carrier, y= arr_delay)) + 
  geom_violin()+
  labs(x='Carrier', y='Distance')

```

