---
title: "week-3-tasks"
author: "Viki-Misi"
date: "October 07, 2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

Produce your first pdf with rmd<br>
install the following than knit a sample rmd file into pdf

```{r, eval=FALSE}
install.packages('tinytex')
tinytex::install_tinytex()
```


Reproduce the followings 

```{r, echo=F}

letters[1:5]
letters[ seq(1,length(letters), by=2) ]

paste('Count',1:10)
paste0(letters[ seq(1,length(letters), by=2) ], collapse = "#")
paste0(letters[ seq(1,length(letters), by=2) ], collapse = " $ ")


```

Print out the value of 

* pi
* round pi to two digits
* squre root of pi
* pi power of 5
```{r, echo=FALSE}
pi
round(pi, 2)
sqrt(pi)
pi^5

```

Create these sequences:
```{r, echo=F}
seq(0, 1, by = 0.02)
seq(1, 2, length.out=13 )
```

* generate a random number between 0 and 1
* generate 10 random number between 0 and 1
* generate 10 random number between 0 and 100
* generate 20 integer random number between 0 and 10 (it should contain 0 and 10 as well and use `runif`)
* generate 20 integer random number between 0 and 10 (use `sample`)



```{r, echo=FALSE}
runif(1)
runif(10)
runif(10)*100
floor(runif(20)*11)
sample(1:10, 20, replace = T)

```


Create a vector called animals
```{r}
animals <- c('dog', 'cat', 'mouse', 'uniconr', 'bear', 'snake', 'whale', 'bull' )

```

* check if 'bear' is in the animals vector
* check if 'cat' and 'snake' are in the animal vector
* sum how many of the followings 'mouse', 'wolf', 'shark', 'bull' are in the animals vector
* how many of the animals start with the letter b


```{r, echo=F}
'bear' %in% animals
c('cat', 'snake') %in% animals
sum(c('mouse', 'wolf', 'shark', 'bull' ) %in% animals)
sum(startsWith(animals, 'b'))
```

Create a list which contains 3 different type of element: string, number and vector
```{r}
l <- list('hello', 10, letters)

```

* print out the second element of the vector
* print out the length of the vector inside the list
* check if the number in the list is divisible by 5 
```{r, echo=F}
l[[3]][2]
print(length(l[[3]]))
l[[2]]%%5 == 0

```



Plot sinus curve from 0, to 2pi
```{r, echo=F}
x <- seq(0, pi*2, by = 0.1)

plot(x, sin(x), type = 'l', col = 'red')

```



Data type/structures

List as many data type structure as you can!

For example:
```{r}
#numeric
i <- 3
str(i)
typeof(i)

# string
s <- 'Hello CEU'
str(s)
typeof(s)

# 
```

