---
title: "week-4-tasks"
author: "Viki-Misi"
date: "October 07, 2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


Install the `nycflights13` package and check it contents.
```{r, message=FALSE}


# install.packages('nycflights13')
library(nycflights13)
library( tidyverse )

```


Read the `my_csvfile.csv` from `https://github.com/CEU-Economics-and-Business/ECBS-5208-Coding-1-Business-Analytics/tree/master/class_2/data/hotels/clean` repository.

```{r, echo=FALSE}

read.csv('https://raw.githubusercontent.com/CEU-Economics-and-Business/ECBS-5208-Coding-1-Business-Analytics/master/class_2/data/hotels/clean/my_csvfile.csv') %>% head()
```


Filter the `flights` data where departure delays was over 1000.
```{r, echo=F}
filter(flights, dep_delay>1000)

```


Filter the `flights` data where departure delays was over 500 and destination was `MIA`  and carrier either `EV, AA, US` .
```{r, echo=F}
filter(flights, dep_delay>500,
       dest=='MIA' & 
         carrier%in% c("EV" ,'AA',"US"))

```


Merge the result with the airlines datasets.

```{r, echo=FALSE}
df <- filter(flights, dep_delay>500,
       dest=='MIA' , 
         carrier%in% c("EV" ,'AA',"US"))

merge(df, airlines, by.x = 'carrier', by.y = 'carrier')


```

Write out the result into

* csv without rownames
* excel
* rds

```{r, echo=FALSE}
write.csv(df,  'df.csv', row.names = F )

library( xlsx )
write.xlsx( df , 'df.xlsx' , sheetName = "Sheet1")

saveRDS(df, 'df.rds')


```





Create a new column in the flights dataset  of the first three column which will be a string like `2015-01-15`
```{r, echo=FALSE}
fl <- flights
fl$my_date <- paste0(fl$year, '-',
                     ifelse(fl$month<10, paste0('0', fl$month), fl$month ), '-',
                     ifelse(fl$day<10, paste0('0', fl$day), fl$day )
)

fl %>% select(c('year', 'month', 'day', 'my_date')) %>% head()

```
Change it to date column
```{r, echo=FALSE}
fl$my_date <- as.Date(fl$my_date)

fl %>% select(c('year', 'month', 'day', 'my_date')) %>% head()

```

Create a new column `delays` which will indicate the delays, sum of absulute values of two delays 

```{r, echo=FALSE}
fl$delays <- abs(fl$dep_delay)+ abs(fl$arr_delay)

fl %>% select(c('year', 'month', 'day', 'dep_delay', 'arr_delay', 'delays')) %>% head()

```


What is the carrier name of the flights which had the most departure delay. 
```{r, echo=FALSE}
merge(flights[which.max(flights$dep_delay),], airlines, by.x = 'carrier', by.y = 'carrier') %>% select('name')



```




