---
title: "week-6"
author: "Viki-Misi"
date: "October 28, 2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r warning=F, message=F, warning=FALSE}
library(nycflights13)
library(tidyverse)
library(ggplot2)
set.seed(24)

```


**Task:**<br>
The best carriers based on arriving (on average the arr_delay is the least), colors= limegreen and cyan4
```{r, echo=FALSE, message=F, warning=FALSE}
# number of flights from origin
flights %>%
  group_by(carrier)%>%
  summarise('mean_arr_delay'=mean(arr_delay, na.rm = T)) %>%
  arrange(mean_arr_delay) %>% 
  ggplot(aes(x = reorder(carrier, mean_arr_delay), y = mean_arr_delay)) + 
    geom_col(color='limegreen', fill='cyan4')+
    labs( y='Average delay', x = 'Carrier')+theme_linedraw()
  


```


**Task:**<br>
Average arrival delay by destinations **name** in August

```{r, echo=FALSE, message=F, warning=FALSE}

flights %>%
  filter(month==8) %>% 
  group_by(dest)%>%
  summarise('mean_delay'=mean(arr_delay)) %>%
  arrange(mean_delay) %>% 
  head(10) %>%
  merge(y = airports %>% select('faa', 'name'), by.x = 'dest', by.y = 'faa' ) %>% 
  arrange(mean_delay) %>% 
  ggplot(aes(x = reorder(name, mean_delay), y = mean_delay)) + geom_col()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.8, hjust=1))+
  labs( y='Average delay', x = 'Destination', title='Average arrival delay', subtitle = 'by destination')



```

**Task:**<br>
The most popular airiles **name** from `JFK`
```{r, echo=FALSE}

flights %>%
  filter(origin=='JFK') %>% 
  group_by(carrier) %>%
  summarise('n_flights'=n()) %>%
  arrange(-n_flights) %>% 
  head(10) %>% 
  merge(y = airlines) %>% 
  ggplot(aes(x = reorder(name, n_flights), y = n_flights)) + geom_col()+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 25, vjust = 0.8, hjust=1))+
  labs( y='Number of flights', x = 'Destination')+
  coord_flip()




```


**Task:**<br>
Write a while loop that prints out standard random normal numbers (use `rnorm()`) but stops (`breaks()`) if you get a number bigger than 1.

```{r, echo=FALSE}
while (TRUE ) {
  t <- rnorm(1)
  if (t>1) {
    print(paste0(t, ' is bigger than 1 the loop stopped'))
    break() 
  }else{
    print(t)
  }
}


```


**Task:**<br>
Adapt the loop from last exercise using `next()` do not print negative numbers.

```{r, echo=FALSE}
while (TRUE ) {
  t <- rnorm(1)
  if (t<0) {
    next()
  }
  if (t>1) {
    print(paste0(t, ' is bigger than 1 the loop stopped'))
    break() 
  }else{
    print(t)
  }
}


```



